---
output:
    word_document:
      fig_caption: true
---

```{r setup, message=FALSE, echo=FALSE}
devtools::load_all()
```

**Short Assignment: NFS484**

*De novo* lipogenesis (DNL) is the process of converting carbohydrates into 
fatty acids. Fatty acids are essentially produced from carbohydrates in this
order: 14:0, 14:1n-7, 16:0, 16:1n-7, 18:0, 18:1n-9 (or 18:1n-7), 20:0, and
20:1n-9. This process of creating fatty acids is important since fatty acids are
necessary for the normal functioning of the body, such as for hormones or 
cellular structures (e.g. cell membrane). However, recent studies suggest that 
excess DNL may be harmful to health. A study was conducted on a cohort of 
participants who are at risk for various chronic diseases. Fatty acids were 
measured along with the Metabolic Syndrome (MetS) components, which is a cluster of 
measures that are risk factors for cardiovascular disease and include blood 
pressure (systolic: SBP, and diastolic: DBP), fasting blood glucose (FPG), HDL 
(1/HDL is the inverse of HDL), triglycerides (TG), and waist size.

```{r plot, fig.cap="Figure: Each dot is represents an association between the MetS component and a fatty acid. The large dot means it is a significant association. The value of the dot on the x-axis is interpreted as a standard deviation (SD) increase in the MetS component for every one SD increase in the fatty acid (for example, for 16:0 with DBP that has a value of 0.1 means that when 16:0 increases by one SD DBP increases by 0.1 SD). A positive value means harmful and a negative value means beneficial.", fig.width=7}
plot_gee_main()
```

**Questions:**

1. Based on your understanding of energy metabolism, individuals consuming a low
glycemic index carbohydrate diet would be at higher risk for production of
excess amounts of the smaller fatty acids (14:0, 16:0, 14:1n-7, 16:1n-7) due to
DNL compared to individuals consuming a high glycemic index diet.
    - True
    - False*

2. Higher 16:0 is associated positively (harmfully) with nearly all aspects of
the MetS
    - True*
    - False

3. All saturated fatty acids are positively (harmfully) associated with MetS
    - True
    - False*

4. Monounsaturated omega-9 fatty acids (18:1n-9 and 20:1n-9) are negatively
(beneficially) associated with HDL and TG
    - True*
    - False

5. More than four fatty acids have positive (harmful) associations with MetS
    - True
    - False*

6. In terms of metabolic health and based on the results of the figure, are
higher amounts of the DNL fatty acids
    - Harmful*
    - Beneficial
    - Neither since both harmful and beneficial fatty acids are produced
    - Unable to tell from the data provided

7. Based on these results, which would be the most likely conclusion if refined
carbohydrates were reduced in a person's diet
    - The person would have no change in the risk of MetS and cardiovascular
    disease
    - Less of the DNL fatty acids would be produced and risk of disease would be
    lower*
    - There would be more DNL fatty acids and metabolic health would be lower
    - Caloric restriction and weight loss would be required before beneficial
    changes in the risk of MetS and cardiovascular disease were observed

8. Based on this assignment and from the tutorial material, individuals who have
a low-to-moderate post meal respiratory quotient (RQ) compared to those with a
high post meal RQ would be at higher risk of accumulating harmful fatty acids
    - True*
    - False

9. With aging, there is an increased prevalence of MetS components and the
tendency for higher fasting RQ. Given this you would predict higher
concentrations of 14:0 and 16:0 fatty acids in an older compared to a younger
population.
    - True*
    - False

10. Risk of excess DNL and developing components of the MetS due to
inappropriate diet, lifestyle and life stage is more likely to be evident in:
    - Post menopausal women consuming a high carbohydrate diet
    - Post menopausal women attempting weight loss through energy restriction
    - Middle aged men and women with central obesity consuming a high
    carbohydrate diet* (?)
    - Middle aged men and women with central obesity attempting weight loss
    through energy restriction
